---
title: "Week08_PortalProject"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.path = "../Outputs/")
```

## Introduction 
This dataset comes from a long-term ecological research site studying the dynamics of desert rodents, plants, ants and weather in Arizona. 
-  This study has produced over 100 scientific papers and is one of the longest running ecological studies in the U.S!

## Load libraries
```{r}
library(tidyverse)
library(here)
library(PNWColors)
library(ggridges)
```

## Read in Data 
```{r}
plots <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-02/plots.csv')
species <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-02/species.csv')
surveys <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-02/surveys.csv')
```

## Explore and Tidy the data 
```{r}
unique(surveys$age)
## taking inspiration from a density plot on twitter by Martijn van Bloois

# going to join the species and surveys datasets 
## look at species on y axis and either year or mean weight on x axis 

spp_surveys <- surveys %>% 
  inner_join(species, 
             by = join_by(species)) %>% 
  filter(! is.na(species)) 
  
```
 
## Plot the Data 
```{r, meanweight}
ggplot(data=spp_surveys, 
       aes(x=meanwgt, 
           y=commonname, 
           fill=meanwgt)) + 
  geom_density_ridges_gradient(scale=2) + 
  scale_fill_viridis_c(option = "D", direction = -1) + 
  labs(x="Meight Weight (g)", 
       y="Common Name of Species", 
       fill="Mean Weight",
       title="Mean Weight by Rodent Species", 
       subtitle=str_wrap("This is a subset of information from an ongoing long-term ecological study that investigates various interactions between organisms and their response to climate change"), ## learning how to include subtitles and captions in the plot! 
       caption = "Data from Portal Project") + 
  theme_classic() + 
  theme(plot.title = element_text(hjust=0.5), 
        axis.title.x = element_text(size=12), 
        axis.title.y = element_text(size=12), 
        axis.text.x=element_text(size=10), 
        axis.text.y=element_text(size=10)) 
ggsave(here("Week08_PortalProject", "Outputs", "meanweight.jpg"), 
       width=10, height=8)
```
## Another plot to visualize the data 
```{r, treatments}
## compare the largest and smallest mouse to see how they survive in treatments across years 

## Need new dataset to organze data 
treatments <- spp_surveys %>% 
  select(treatment, commonname, year, plot) %>% 
  filter(commonname=="Banner-tailed kangaroo rat"|commonname=="Silky pocket mouse", 
         treatment=="removal"|treatment=="control") %>% 
  mutate(year=factor(year)) %>% 
  group_by(year, commonname, treatment) %>% 
  summarize(total=sum(plot))

ggplot(data=treatments, 
       aes(x=year, 
           y=total, 
           fill=commonname)) + 
  geom_col() + 
  facet_wrap(~treatment) + 
  scale_fill_manual(values=pnw_palette("Starfish", n=2)) + 
  theme_bw() + 
  labs(x="Year", 
       y="Total counts across plots", 
       fill="Common Name", 
       title="Comparison of Kangaroo Rat and Pocket Mouse between treatments from 1978-2018", 
       caption="Banner-tailed kangaroo rat does better in control treatment plots than silky pocket mouse, which thrives more in removal treatments but in later years") + 
  theme(plot.title = element_text(hjust=0.5), 
        axis.title.x = element_text(size=12), 
        axis.title.y = element_text(size=12), 
        axis.text.y=element_text(size=10), 
        axis.text.x=element_text(size=10, angle = 45, vjust=0.5)) + 
  scale_x_discrete(breaks = c("1978", "1985", "1990", "1995", "2000", "2001", "2005", "2010", "2015"))
ggsave(here("Week08_PortalProject", "Outputs", "treatments.jpg"), 
       width=12, height=8)
```

