---
title: "TidyTuesday05_Drugs"
author: "Hannah Merges"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: tactile
    toc: TRUE
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message=FALSE, 
                      fig.path = "../Outputs/")
```
## Introduction  
This week's dataset includes information about drug development and approval for both human and animal patients for various different conditions.  
There are many ways to analyze all the information included in here so let's first look at what is in each column...

## Load libraries 
```{r, echo=FALSE}
??ggsave
library(tidyverse)
library(here)
library(praise)
library(beyonce)
library(lubridate)
library(ggthemes)
??ggthemes()
library(devtools)
library(patchwork)
library(ggsn)
library(gridExtra)
library(ggridges)
library(hrbrthemes)
library(janitor)
??janitor

```

## Load the Data 
```{r}
drugs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-03-14/drugs.csv')
```
## Look at the data, tidy/wrangle it  
```{r}
distinctcolumns <- drugs %>% 
  distinct(medicine_name)
## is there a way to see how many times a specific ID was used? YES --> see below 
active_subs <- drugs %>% 
  count(active_substance) %>% 
  arrange(desc(n)) %>% 
  slice_head(n = 100)

####checking with medicine names 
medicines <- drugs %>% 
  count(medicine_name) %>% 
  arrange(desc(n))

##learned this from a new code -- create another df with the counts for each of the therapeutic area 
number_theraparea <- drugs %>% 
  count(therapeutic_area) %>% ##this gives you the counts for each individual therapeutic area 
  arrange(desc(n)) %>% ##this arranges them in descending order - starting with highest count 
  drop_na() 
  #slice_head(n = 100) ##this would cut it at whatever the top 100 are, but I want all of them so I will not use it in my code

#now you can view both of these together, expanded, using expand_grid()
full_drug_grid <- expand_grid(active_substance = active_subs$active_substance, 
                         therapeutic_area = number_theraparea$therapeutic_area) ##this expands both columns so that for each therapeutic area, we can see which substances are active in the drugs used for that area

#join these two dfs together to see how each therap. area matches with the active drugs 
joined_drug_grid <- drugs %>% 
  semi_join(active_subs, by = "active_substance") %>% 
  semi_join(number_theraparea, by = "therapeutic_area") %>% 
  count(active_substance, therapeutic_area) %>% ##adds a column to show counts for the active sub per therapeutic area 
  filter(therapeutic_area == "Hypertension") 

##trying a new way to look at the data because I am struggling to create the plots in a meaningful way
joinedgrid2 <- full_join(medicines, drugs) %>% 
  filter(therapeutic_area == "Hypertension") %>% 
  select(1,2,4,5,6,9)

#somehow need to get it so that the active ingredients are not separated by commas ??
#couldn't figure out how to get this to work so just left it with full list of active substances
##df_base <- joined_drug_grid %>% 
  #left_join(number_theraparea,active_subs, by = c("therapeutic_area")) %>% 
  #mutate(n = replace_na(n, 0)) %>% 
  #arrange(desc(n), active_substance, therapeutic_area) %>% 
  #mutate(#x = as.numeric(factor(active_substance)),
          #y = as.numeric(factor(therapeutic_area))) %>% 
  #filter(n > 0)

#as.numeric(joined_drug_grid$n)
###oooo it is not working because there are multiple names in this column ??

```
## Time to make a plot! 
Going to look at the active substances in medicines commonly used for hypertension.
```{r, drugs, fig.cap="The list of most common active ingredients and subsntances in hypoertension medication and how often each of these active substances is used per type of medication."}

ggplot(data=joined_drug_grid, 
       aes(x=active_substance, 
           y=n,
           fill=active_substance)) + 
  geom_col() +
  labs(x="Active substances", 
       y="Frequency of active substances", 
       title="Frequency of active substances found in different hypertension medications", 
       fill="Active Substances") + 
  theme_bw() +
  coord_flip() + 
  theme(plot.title=element_text(h=0.5))
  
ggsave(here("Week05_Drugs","Outputs","hypertenionplot.jpg"), 
       width=10, height=10)

```
Have these active substances been withdrawn at all? 
```{r}
ggplot(data=joinedgrid2, 
       aes(x=))
```


## Summary 
This week I learned how to manipulate data in new ways. Something I am struggling with still is how to best organize and manipulate/tidy the data to really get what I want from the plot. Plotting has become easier but now I am realizing I need to arrange the data better. 
Thank you to [Dan Oehm](https://github.com/doehm) for the accessible code to help me learn how to group my count data and summarize the columns I wanted to use for plotting this information. 






